\speculake: Hiding Malware in Speculative Execution

### 2018-08-07 paper/architecture.tex
fix citation
typos

Eventually, the dereference of the uncached function pointer in the payload program
will be resolved, and the CPU will recognize it has incorrectly predicted the
destination of its \texttt{callq} instruction. The
results from the speculative entry point instructions will be discarded, and the
CPU will continue executing from the correct destination. However, the
speculative code can change what is loaded into the cache based on its
computation, allowing it to covertly communicate its results to the ``real world'' program.

@@

We found that the processor can identify some operations that have no effect on
output registers and allows them to use the entire size provided by the ROB, as
determined in Section~\ref{sssec:ROB}. This includes idomatic no-ops and zero
idioms, which have no reliance on register values, for example the instruction
\texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
instruction is a potential data hazard, it must use an entry in the Physical 
Register File (PRF) until it reaches the correct stage in the execution pipeline. 
In this case, the number of instructions could be limited by the PRF
entries available.

@@

Most notably instructions that use the extended x86 registers are still valid
within the speculative context. Specifically, Intel's hardware accelerated
AES-NI encryption and decryption instructions, which each use 128-bit registers.
As shown in Figure~\ref{fig:spec-capacity-np}, speculative environments can
complete a significant number of AES rounds---over 100 rounds in our
experiments, more than enough to decrypt a
full block using simple AES modes (e.g. AES-CTR). We investigate the use of AES
instructions in the spculative environment further in
Sections~\ref{subsec:decryption}.

@@

We find that when executing speculatively, the number of instrutions completed
is rarely limited by cache miss duration. Instead
ROB size and processor mechanisms for resolving
data dependencies define an upper boundary of approximately 150
instructions\footnote{While \texttt{nop} is able to execute up to the full 220
ROB capacity, instructions that do useful work (and/or use multiple $\mu$ ops) cannot reach this limit}.

@@

We find that not all Intel CPUs support nested speculation. For
example, it appears Haswell chips do not speculate while already speculating,
but nonetheless support non-nested \speculake. Both Sandy Bridge (which
preceeded Haswell) and Kaby Lake (which followed Haswell) support nested
speculation. We find that when a CPU does support nested speculation, there
appears to be no limit to nested depth besides the speculative instruction
limit. We use a 16-deep nested speculation in Section~\ref{subsec:impl-aes}
to protect speculative decryption keys from reverse
engineering.

@@

We summarize our findings into a \emph{speculative primitive}, which allows our payload program to
speculatively (and covertly) perform on the order of 100 arbitrary
instructions while an accompanying trigger program is running, and communicate
a short (e.g. single byte) result to the real
world via a cache side channel. These speculative instructions are able to read
from any real-world memory or registers, but they cannot perform updates or
writes directly. To update memory, the speculative instructions must communicate
to the real world. We use a cache side channel to do so, but other side channels
compatible with Spectre could also be used~\cite{spectre1-1}.

>>>>

Eventually, the de-reference of the uncached function pointer in the payload
program will be resolved, and the CPU will recognize it has incorrectly
predicted the destination of its \texttt{callq} instruction. The results from
the speculative entry point instructions will be discarded, and the CPU will
continue executing from the correct destination. However, the speculative code
can change what is loaded into the cache based on its computation, allowing it
to covertly communicate its results to the ``real world'' program.

@@

We found that the processor can identify some operations that have no effect on
output registers and allows them to use the entire size provided by the ROB, as
determined in Section~\ref{sssec:ROB}. This includes idiomatic no-ops and zero
idioms, which have no reliance on register values, for example the instruction
\texttt{xchg \%rax, \%rax} and \texttt{xor \%rax, \%rax}. However, if an
instruction is a potential data hazard, it must use an entry in the Physical 
Register File (PRF) until it reaches the correct stage in the execution pipeline. 
In this case, the number of instructions could be limited by the PRF
entries available.

@@

Most notably instructions that use the extended x86 registers are still valid
within the speculative context. Specifically, Intel's hardware accelerated
AES-NI encryption and decryption instructions, which each use 128-bit registers.
As shown in Figure~\ref{fig:spec-capacity-np}, speculative environments can
complete a significant number of AES rounds---over 100 rounds in our
experiments, more than enough to decrypt a
full block using simple AES modes (e.g. AES-CTR). We investigate the use of AES
instructions in the speculative environment further in
Sections~\ref{subsec:decryption}.

@@

We find that when executing speculatively, the number of instructions completed
is rarely limited by cache miss duration. Instead
ROB size and processor mechanisms for resolving
data dependencies define an upper boundary of approximately 150
instructions\footnote{While \texttt{nop} is able to execute up to the full 220
ROB capacity, instructions that do useful work (and/or use multiple $\mu$ ops) cannot reach this limit}.

@@

We find that not all Intel CPUs support nested speculation. For
example, it appears Haswell chips do not speculate while already speculating,
but nonetheless support non-nested \speculake. Both Sandy Bridge (which
preceded Haswell) and Kaby Lake (which followed Haswell) support nested
speculation. We find that when a CPU does support nested speculation, there
appears to be no limit to nested depth besides the speculative instruction
limit. We use a 16-deep nested speculation in Section~\ref{subsec:impl-aes}
to protect speculative decryption keys from reverse
engineering.

@@

We summarize our findings into a \emph{speculative primitive}, which allows our
payload program to speculatively (and covertly) perform on the order of 100
arbitrary instructions while an accompanying trigger program is running, and
communicate a short (e.g. single byte) result to the real world via a cache side
channel. These speculative instructions are able to read from any real-world
memory or registers, but they cannot perform updates or writes directly. To
update memory, the speculative instructions must communicate to the real world.
We use a cache side channel to do so, but other side channels compatible with
Spectre could also be used~\cite{kiriansky2018speculative}.

### 2018-08-07 paper/figures.tex
typos

\newcommand{\FigSpecMeasureParity}{
\begin{figure}[b]
     \centering
        \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
        \caption{\textbf{Hyperthreading}\,---\,%
        The trigger program influences the payload program via the indirect branch predictor
        when running on the same logical CPU core. Hyperthreads also share an indirect branch
        predictor despite running on different logical cores. We measured the impact of
        same logical core (warm colors) vs paritiy hyperthreads (cool colors, denoted by (P))
        on the size of the speculative execution window. As hyperthreads run concurrently, they do not
        need to wait for an OS context switch between processes, but the two processes directly
        contend for ROB entries. As a result, we find that hyperthreads generally have a smaller
        speculative execution window.}
    \label{fig:spec-capacity-parity}
\end{figure}
}

@@

\newcommand{\FigSpasmModel}{
\begin{figure}[b]
    \centering
        \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
        \caption{\textbf{SPASM model}\,---\,%
        Our SPASM emulator speculatively decrypts intructions, and emulates
        them in the real world. The \textit{Speculative Computation} decrypts
        the encrypted SPASM binary, returning the result through the side channel
        to allow the \textit{Resultant Computation} to update the emulated state
        (or make system calls) on behalf of the speculative world.}
    \label{fig:spasm_model}
\end{figure}
}

>>>>

\newcommand{\FigSpecMeasureParity}{
\begin{figure}[b]
     \centering
        \includegraphics[width=1.05\linewidth]{figures/Speculative_measurements_parity.png}
        \caption{\textbf{Hyperthreading}\,---\,% The trigger program influences
        the payload program via the indirect branch predictor when running on
        the same logical CPU core. Hyperthreads also share an indirect branch
        predictor despite running on different logical cores. We measured the
        impact of same logical core (warm colors) vs parity hyperthreads (cool
        colors, denoted by (P)) on the size of the speculative execution window.
        As hyperthreads run concurrently, they do not need to wait for an OS
        context switch between processes, but the two processes directly contend
        for ROB entries. As a result, we find that hyperthreads generally have a
        smaller speculative execution window.}
    \label{fig:spec-capacity-parity}
\end{figure}
}

@@

\newcommand{\FigSpasmModel}{
\begin{figure}[b]
    \centering
        \includegraphics[width=0.46\textwidth]{figures/spasm_model.pdf}
        \caption{\textbf{SPASM model}\,---\,%
        Our SPASM emulator speculatively decrypts instructions, and emulates
        them in the real world. The \textit{Speculative Computation} decrypts
        the encrypted SPASM binary, returning the result through the side channel
        to allow the \textit{Resultant Computation} to update the emulated state
        (or make system calls) on behalf of the speculative world.}
    \label{fig:spasm_model}
\end{figure}
}

### 2018-08-07 paper/implementation.tex
typos

The local array store is now the vulnerability it can be exploited even 
with complete bounds checks. After a series of honest inputs we include 
an entry with a padding length beyond the maximum. On the runs with this
cafted input we flush from cache the maximum padding length constant in 
the payload program, preventing the bounds check from being resolved 
and allowing the data value to be speculatively stored. We 
then speculatively overwrite the return address and use store-to-load forwarding
to cause a following return to divert to the address specified by the data entry. 
From here we make use of the decryption gadget to dcecrypt SPASM instructions
from an encrypted data blob to be transmitted to the real world via cache 
side-channel. Using this method we have been able to run all SPASM payloads 
including the reverse-shell.

>>>>

The local array store is now the vulnerability it can be exploited even with
complete bounds checks. After a series of honest inputs we include an entry with
a padding length beyond the maximum. On the runs with this crafted input we
flush from cache the maximum padding length constant in the payload program,
preventing the bounds check from being resolved and allowing the data value to
be speculatively stored. We then speculatively overwrite the return address and
use store-to-load forwarding to cause a following return to divert to the
address specified by the data entry. From here we make use of the decryption
gadget to decrypt SPASM instructions from an encrypted data blob to be
transmitted to the real world via cache side-channel. Using this method we have
been able to run all SPASM payloads including the reverse-shell.

### 2018-08-07 paper/triggers.tex
typos

To do this, we use Specualtive Buffer Overflows
(SBO)~\cite{kiriansky2018speculative} to redirect control flow to a speculative
entry point specified by user input. We design a payload program that takes
arbitrary user input and performs appropriate bounds checks to ensure no
traditional control flow violations could be exploited. However, using the
Spectre~1.1 variant, control flow can still be violated speculatively, allowing
the adversary to force a speculative entry point based entirely off the input
provided. This allows the trigger to be an input, potentially even provided over
a network if the payload program accepts network data. Traditional symbolic
execution and code coverage fuzzers will be unable to discover this trigger
input, as they do not model the speculative state of the CPU.

>>>>

To do this, we use Speculative Buffer Overflows
(SBO)~\cite{kiriansky2018speculative} to redirect control flow to a speculative
entry point specified by user input. We design a payload program that takes
arbitrary user input and performs appropriate bounds checks to ensure no
traditional control flow violations could be exploited. However, using the
Spectre~1.1 variant, control flow can still be violated speculatively, allowing
the adversary to force a speculative entry point based entirely off the input
provided. This allows the trigger to be an input, potentially even provided over
a network if the payload program accepts network data. Traditional symbolic
execution and code coverage fuzzers will be unable to discover this trigger
input, as they do not model the speculative state of the CPU.
